---
title: "week6"
author: "Zach Schroeder"
date: "5/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practice models

```{r}
library(tidyverse)
library(lme4)
d <- read_csv(here::here("data", "longitudinal-sim.csv"))
m <- lmer(g5_spring ~  (1|scid) + (1|distid) +g4_spring + g3_spring, d)

#Exercise 2
d <- d %>% 
  group_by(scid) %>% 
  mutate(sch_mean_start = mean(g3_fall)) %>% 
  group_by(distid) %>% 
  mutate(dist_mean_start = mean(g3_fall))
m <- lmer(g5_spring ~ g4_spring + g3_spring + sch_mean_start + (g3_spring|scid) + (1|distid), d)

#Exercise 3
lmer(g5_spring ~ g4_spring + g3_spring + sch_mean_start + 
       (g4_spring + g3_spring|scid) + 
       (g4_spring + g3_spring|distid),
     data = d)

#Exercise 4
lmer(g5_spring ~ g4_spring + g3_spring*sch_mean_start + 
       (g4_spring + g3_spring|scid) + 
       (g4_spring + g3_spring|distid),
     data = d)
lmer(g5_spring ~ g4_spring + g3_spring + dist_mean_start + 
       (g4_spring + g3_spring||scid) + 
       (g4_spring + g3_spring||distid),
     data = d)

# MOve to long
l <- d %>% 
  pivot_longer(
    cols = starts_with("g"),
    names_to = "timepoint",
    values_to = "score"
  )
l
wave_frame <- tibble(
  timepoint = paste0(
    "g", 
    rep(3:5, each = 3), 
    rep(c("_fall", "_winter", "_spring"), 3)
  ),
  wave = 0:8
)
wave_frame
l <- left_join(l, wave_frame)
# Model 6
lmer(score ~ wave +
       (wave|sid) + (wave|scid) + (1|distid),
     data = d)

# Model 7
lmer(score ~ wave*dist_mean_start*sch_mean_start + 
       (wave|sid) + (wave|scid) + (sch_mean_start|distid),
     data = d)
# Model 8
lmer(score ~ sch_mean_start + dist_mean_start + (wave|sid), l)
```

DATA
```{r}
d <- read_csv(here::here("data", "cnlsy.csv"))
library(lme4)
d <- d %>% 
  mutate(wave_c = wave - 1)
m_wave <- lmer(piat ~ wave_c + (wave_c|id),
               data = d)
m_agegrp <- lmer(piat ~ agegrp + (agegrp|id),
                 data = d,
                 control = lmerControl(optimizer = "bobyqa"))
```

Ah ha! We must center!

```{r}
d <- d %>% 
  mutate(agegrp_c = agegrp - 6.5,
         age_c = age - 6)
m_agegrp2 <- lmer(piat ~ agegrp_c + (agegrp_c|id),
                 data = d,
                 control = lmerControl(optimizer = "bobyqa"))
m_age <- lmer(piat ~ age_c + (age_c|id),
                 data = d,
                 control = lmerControl(optimizer = "bobyqa"))
performance::compare_performance(m_agegrp2, m_age,
                    metrics = c("AIC", "BIC"),
                    rank = TRUE) %>% 
  as_tibble()
```

Modeling monthly growth (as opposed to yearly growth)

```{r}
d <- d %>% 
  mutate( age_m = (age - 6)*12)
m_age <- lmer(piat ~ age_m + (age_m|id),
                 data = d,
                 control = lmerControl(optimizer = "bobyqa"))
arm::display(m_age)
```

Wages Data

```{r}
wages <- read_csv(here::here("data", "wages.csv"))
m_wage0 <- lmer(lnw ~ exper + (exper|id), data = wages,
                control = lmerControl(optimizer = "bobyqa"))
arm::display(m_wage0)
```

Is the relation between experience and log wages the same across coded race/ethnicity categories? Do these relations depend on highest grade completed?

```{r}
wages <- wages %>% 
  mutate(hgc_9 = hgc-9)
wage1 <- lmer(lnw ~ exper + hispanic + black + hgc_9 + (exper|id), 
              data = wages,
              control = lmerControl(optimizer = "bobyqa"))
```

Plotting

```{r}
m_wage1 <- lmer(lnw ~ exper + black + hispanic + hgc_9 +
                  (exper|id),
                data = wages,
                control = lmerControl(optimizer = "bobyqa"))
pred_frame <- expand.grid(
  exper = 0:15,
  black = 0:1,
  hispanic = 0:1,
  hgc_9 = 6:12 - 9,
  id = -999
)
pred_frame <- pred_frame %>% 
  mutate(pred = predict(m_wage1, 
                        pred_frame, 
                        allow.new.levels = TRUE))
pred_frame <- pred_frame %>% 
  mutate(
    race_eth = case_when(
      black == 0 & hispanic == 0 ~ "White",
      black == 1 & hispanic == 0 ~ "Black",
      black == 0 & hispanic == 1 ~ "Hispanic",
      TRUE ~ NA_character_
    )
  )
pred_frame %>% 
  drop_na() %>% 
  filter(hgc_9 == 0) %>% 
  ggplot(aes(exper, pred)) +
  geom_line(aes(color = race_eth))

m_wage2 <- lmer(lnw ~ exper + black + exper:black +
                  exper:hispanic + 
                  hgc_9 + exper:hgc_9 +
                  (exper|id),
                data = wages,
                control = lmerControl(optimizer = "bobyqa"))
pred_frame <- pred_frame %>% 
  mutate(pred_int = predict(m_wage2, 
                            newdata = pred_frame, 
                            allow.new.levels = TRUE))
pred_frame %>% 
  drop_na() %>% 
  filter(hgc_9 == -3 | hgc_9 == 3) %>% 
  ggplot(aes(exper, pred_int)) +
  geom_line(aes(color = race_eth)) +
  facet_wrap(~hgc_9)
pred_frame %>% 
  drop_na() %>% 
  ggplot(aes(exper, pred_int)) +
  geom_line(aes(color = factor(hgc_9))) +
  facet_wrap(~race_eth) +
  scale_color_brewer("Highest grade completed", 
                     palette = "Accent", 
                     breaks = 3:-3,
                     labels = 12:6) +
  labs(x = "Experience (years)",
       y = "Model Predicted wages (log scaled)")
```

Non linearity

```{r}
sim_d <- read_csv(here::here("data", "curvilinear-sim.csv"))
sim_d %>% 
  arrange(sid, date) %>% 
  group_by(sid) %>% 
  slice(1) %>% 
  ungroup() %>% 
  summarize(range(date))
ggplot(sim_d, aes(date, score)) +
  geom_point(alpha = 0.15, stroke = NA) +
  geom_smooth(se = FALSE, color = "#33B1AE", size = 2)
```

Creating Models
```{r}
sim_d <- sim_d %>% 
  mutate(
    days_from_start = as.numeric(date) - min(as.numeric(date))
  )
linear <- lmer(score ~ days_from_start + (days_from_start|sid), 
               data = sim_d,
               control = lmerControl(optimizer = "Nelder_Mead"))
arm::display(linear)
sim_d <- sim_d %>% 
  mutate(days2 = days_from_start^2)
quad <- lmer(score ~ days_from_start + days2
               (days_from_start|sid), 
             data = sim_d,
             control = lmerControl(optimizer = "Nelder_Mead"))
arm::display(quad)
performance::compare_performance(linear, quad)

sim_d <- sim_d %>% 
  mutate(days3 = days_from_start^3)
cubic <- lmer(score ~ days_from_start + days2 + days3 + 
               (days_from_start|sid), 
             data = sim_d,
             control = lmerControl(optimizer = "Nelder_Mead"))
arm::display(cubic)
performance::compare_performance(linear, quad, cubic)
```

You can also use poly() as a function on days from start and then a comma
cubic <- lmer(score ~ poly(days_from_start, 3) +  
               (days_from_start|sid), 
             data = sim_d,
             control = lmerControl(optimizer = "Nelder_Mead"))
HOWEVER, the base polynomials will have orthogonal polynomials, which means our output will be even less interpretable than the by-hand version above

```{r log growth}
sim_d <- sim_d %>% 
  mutate(days_log = log(days_from_start + 1))
log_m <- lmer(score ~ days_log + (days_log|sid), 
              data = sim_d)
arm::display(log_m)
```

