---
title: "Week10"
author: "Zach Schroeder"
date: "6/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
alc <- read_csv(here::here("data", "alcohol-adolescents.csv"))
```

# Theoretical Models to Code

>Does alcohol use increase with age?

```{r}
library(lme4)
alc <- alc %>% 
  mutate(age_c = age - 14)
rq1 <- lmer(alcuse ~ age_c + (1 + age_c | id),
             data = alc)
summary(alc)
```

> Do children of alcoholics have higher levels of alcohol use?

```{r}
alc <- alc %>% 
  mutate(coa = factor(coa),
         male = factor(male))
rq2 <- lmer(alcuse ~ age_c + coa + (1 + age_c | id),
             data = alc)
```

>Do male children of alcoholics have higher alcohol use than female children of alcoholics?

```{r}
rq3 <- lmer(alcuse ~ age_c + coa*male +
              (age_c | id),
            data = alc)
```

>Do adolescents' alcohol use trajectories differ between males and females?

```{r}
rq4 <- lmer(alcuse ~ age_c*male +
              (age_c | id) ,
            data = alc)
```

>Does the trajectory of alcohol use differ for male children of alcoholics than female children of alcoholics?

```{r}
rq5 <- lmer(alcuse ~ age_c*male*coa +
              (age_c | id) ,
            data = alc)
```

```{r}
threelev <- read_csv(here::here("data", "three-lev.csv"))
threelev
```

>To what extent do math scores, and changes in math scores, vary between students versus between schools?

```{r}
rq1_math <- lmer(math ~ 1 + year + 
                   (year | sid) + (year | schid),
                 data = threelev)

as.data.frame(VarCorr(rq1_math))

as.data.frame(VarCorr(rq1_math)) %>% 
  mutate(prop = vcov / sum(vcov))

as.data.frame(VarCorr(rq1_math)) %>% 
  mutate(prop = vcov / sum(vcov)) %>% 
  ggplot(aes(grp, prop)) +
  geom_col(aes(fill = var1))
```

>What is the average difference in math scores among students coded Black or Hispanic, versus those who are not?

```{r}
rq2_math <- lmer(math ~ year + black + hispanic +
                   (year | sid) + (year | schid),
                 data = threelev,
                 control = lmerControl(optimizer = "bobyqa"))
```

> To what extent do the differences in gains between students coded Black or Hispanic, versus those who are not, vary between schools?

```{r}
rq3_math <- lmer(math ~ year*black + year*hispanic +
                   (year | sid) + (year*black*hispanic | schid),
                 data = threelev,
                 control = lmerControl(optimizer = "bobyqa"))

as.data.frame(VarCorr(rq3_math)) %>% 
  mutate(prop = vcov / sum(vcov)) %>% 
  ggplot(aes(grp, prop)) +
  geom_col(aes(fill = var1))
```

# Lung cancer remission model

```{r}
library(brms)
lc <- brm(
  remission ~ age * tumorsize + lungcapacity + (1|did),
  data = hdp,
  family = bernoulli(link = "logit"),
  cores = 4,
  backend = "cmdstan"
)
```

```{r}
library(mice) # Missing Data
naniar::vis_miss(nhanes)
mi_nhanes <- mice(nhanes, m = 5, print = FALSE)
m_mice <- brms::brm_multiple(bmi ~ age * chl, 
                       data = mi_nhanes)
```

```{r}
library(lme4)
Penicillin <- Penicillin %>% 
  arrange(sample)
head(Penicillin)
nested1 <- lmer(diameter ~ 1 + (1 | sample), 
                data = Penicillin)
library(sundry)
pull_residual_vcov(nested1) %>% 
  image()
Penicillin <- Penicillin %>% 
  arrange(plate)
crossed1 <- lmer(diameter ~ 1 + (1 | sample) + (1|plate), 
                 data = Penicillin)
pull_residual_vcov(crossed1) %>% 
  image()
```

```{r}
data(Oxide, package = "nlme")
head(Oxide)
nested2 <- lmer(Thickness ~ 1 + (1|Lot/Wafer), data = Oxide)
pull_residual_vcov(nested2) %>% 
  image()
crossed2 <- lmer(Thickness ~ 1 + (1|Lot) + (1|Wafer), 
                 data = Oxide)
pull_residual_vcov(crossed2) %>% 
  image()
```

```{r}
achievement <- read_csv(here::here("data", "pupils.csv"))
achievement
ccrem <- lmer(achievement ~ 1 + 
                (1|primary_school_id) + (1|secondary_school_id),
              data = achievement)
arm::display(ccrem)
str(ranef(ccrem))
pull_residual_vcov(ccrem) %>% 
  image()
```

```{r}
mm_brms <- brm(
  achievement ~ 1 + 
    (1 | mm(primary_school_id, secondary_school_id)),
  data = achievement
)
summary(mm_brms)
```

```{r}
set.seed(42)
achievement <- achievement %>% 
  mutate(primary_school_time = abs(rnorm(1000)),
         secondary_school_time = abs(rnorm(1000))) %>% 
  rowwise() %>% 
  mutate(tot = primary_school_time + secondary_school_time,
         primary_school_time = primary_school_time / tot,
         secondary_school_time = secondary_school_time / tot) %>% 
  ungroup()
achievement %>% 
  select(PUPIL, ends_with("time"))

mm_brms2 <- brm(
  achievement ~ 1 + 
    (1 | mm(primary_school_id, secondary_school_id,
            weights = cbind(primary_school_time, 
                            secondary_school_time))),
  data = achievement)
```

